\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{url}
\title{????? PFAS?? ??????-???????? ????? ?????(2-??????? ?????? root??? ?????????? Vmax ????)??? ??????????}
\author{sci-adk (deterministic render)}
\date{\today}

\begin{document}
\maketitle

\noindent\textit{Draft compiled by sci-adk from Spec \texttt{pfas-rice-carrier} (v1). Belief state is revisable as Evidence accrues.}

\section{Goal}
????? PFAS?? ??????-???????? ????? ?????(2-??????? ?????? root??? ?????????? Vmax ????)??? ??????????
n\_C?? ????????? ?????????? ????????? QSPR?? ????? ????????? (?????? ?????? R\textasciicircum{}2 > 0.9????? ????????,
?????????? ad-hoc).

\section{Background}
?????(C10--C12) PFAS?? ????? ???????????? LC1--LC5???? ???????????? ??????????: ?????????
???????????? ????? throttle(LC1), ????????? ?????-?????? ?????? ?????(LC2), ??????-?????? root???
?????(LC3), 2-?????? C10--C11??? ??????(LC4), PFDoDA?? uptake ???????? ??? ????????? ??????
?????? ???????? ??????(LC5). ?????? LC5b?? PFDoDA?? ??? Vmax \textasciitilde{}5$\times$??? ??? ??????????? illustrative
??????????. ?? ???????? ??????? ???????????? ????????? ?????????? ????????(QSPR) ????????,
??????????? ad-hoc ??????? ?????????.

\section{Method}
2-??? ???????????(`validation/twopool\_longchain.py`, ORYZA2000 biomass)???? ??? ?????
????????(PFOA$\cdot$PFNA$\cdot$PFDA$\cdot$PFUnDA$\cdot$PFDoDA)?? ???? g\_xy=g\_ph=0(uptake$\rightarrow$root ?????)???? ????
Yamazaki ?????? root??? ?????????? Vmax ??????? brentq?? ???????????. log10(Vmax ????)??? ??????????
n\_C ??? ??? ????????? log10 K\_PL?? ?????????????? ????????? R\textasciicircum{}2??? ???????????. ??????(Yamazaki root)??
????? in-sample ??????????? ??????????.

Planned approaches:
\begin{itemize}
  \item 2-??? ???????????(`validation/twopool\_longchain.py`, ORYZA2000 biomass)???? ??? ?????
????????(PFOA$\cdot$PFNA$\cdot$PFDA$\cdot$PFUnDA$\cdot$PFDoDA)?? ???? g\_xy=g\_ph=0(uptake$\rightarrow$root ?????)???? ????
Yamazaki ?????? root??? ?????????? Vmax ??????? brentq?? ???????????
  \item log10(Vmax ????)??? ??????????
n\_C ??? ??? ????????? log10 K\_PL?? ?????????????? ????????? R\textasciicircum{}2??? ???????????
  \item ??????(Yamazaki root)??
????? in-sample ??????????? ??????????.
\end{itemize}

\section{Hypotheses and findings}

\subsection{????? PFAS?? ??????-???????? ????? ?????(2-??????? ?????? root??? ?????????? Vmax ????)??? ??????????
n\_C?? ????????? ?????????? ????????? QSPR?? ????? ????????? (?????? ?????? R\textasciicircum{}2 > 0.9????? ????????,
?????????? ad-hoc).}
\begin{itemize}
  \item Hypothesis id: \texttt{hyp-001} (exploratory)
  \item Decision rule (qualitative): Expert judgment based on evidence
  \item \textbf{Status: refuted} --- confidence strong (graded)
  \item Basis: qualitative judge: The carrier-enhancement factor is NOT a smooth log-linear function of chain length (R\textasciicircum{}2 0.70 on n\_C, 0.62 on K\_PL; both <0.9): essentially no enhancement is needed to C10, then a steep onset at C11-C12. So the long-chain active-carrier enhancement is NOT cleanly QSPR-able -- it stays a longest-chain-specific (ad-hoc) lever, not a chain-length-predictable one. Hypothesis refuted.
  \item \textit{Evidence validity:} referent=empirical; data\_source(s)=measured
\end{itemize}

\section{Evidence}
\begin{itemize}
  \item \texttt{evi-carrier-qspr} (experiment\_run): point=0.701, finding=LC6: Vmax multiplier reproducing measured ROOT (g\_xy=g\_ph=0) per congener -- PFOA 1.2x, PFNA 1.3x, PFDA 1.2x, PFUnDA 2.0
\end{itemize}

\section{Pending agent judgments}
These hypotheses use a non-numeric rule and await an in-session agent verdict (no autonomous LLM call):
\begin{itemize}
  \item \texttt{hyp-001} (qualitative): Expert judgment based on evidence
  \item finding: LC6: Vmax multiplier reproducing measured ROOT (g\_xy=g\_ph=0) per congener -- PFOA 1.2x, PFNA 1.3x, PFDA 1.2x, PFUnDA 2.0x, PFDoDA 5.5x. Regression of log10(multiplier): on n\_C slope 0.15/CF2 R\textasciicircum{}2 0.701; on log10 K\_PL slope 0.37 R\textasciicircum{}2 0.620. NOT log-linear (R\textasciicircum{}2<0.9): \textasciitilde{}no enhancement to C10, then a steep threshold-like rise at C11-C12. => the long-chain carrier enhancement is NOT cleanly QSPR-able from chain length; it is a steep onset at the longest chains.
\end{itemize}

\section{References}
No literature cited.

\end{document}