\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{url}
\title{sci-adk?? ????? ???????????? ?????????? ???????????. ?????? ?????? ???????? ?????:}
\author{sci-adk (deterministic render)}
\date{\today}

\begin{document}
\maketitle

\noindent\textit{Draft compiled by sci-adk from Spec \texttt{pfas-rice-longchain} (v1). Belief state is revisable as Evidence accrues.}

\section{Goal}
sci-adk?? ????? ???????????? ?????????? ???????????. ?????? ?????? ???????? ?????:

????? LC1. ???????????? ???????? ?????(C10--C12) shoot ????????? ??????????? ?????????????? ---
f\_xy=1 ???????????? ????? ?????????? Cw=C/B ??????? ????? shoot?? ???????? ????????? ????????.

????? LC2. B-??????? ?????-?????? ?????? ????? ???(g\_xy$\cdot$C, g\_ph$\cdot$C)??? ???? ??????? ????? ????????
????? shoot(straw+grain) ??????? ?????????? ????????.

????? LC3. ?? ????? ???????????? root?? ???$\cdot$????? ????????? ??????????? ????? ????? shoot???
??????????(?????? ??????????? ???????? ?????????).

\section{Background}
PFAS-?? ???????? ?????\textasciitilde{}????? ?????????? ????? BAF??? ????????? ????????????, **?????(C10--C12)
??? ??????????? ?????????**?????. ORYZA2000 biomass???? per-congener ??????????? ????
PFDoDA(C12)??? f\_xy=1$\cdot$L\_Ph=1$\cdot$kappa\_d ???????????????? straw 14.6 vs 49.8, grain 7.4 vs
45.5?? ????????? ????????. ?????? ???????? **????????????(free-anion) ??????? ??????? throttle**
????: ??????$\cdot$????? ????? ??????????? ???? ?????????? Cw=C/B?? ???????????, ?????? B??? ?????
??????? ???? \textasciitilde{}10\textasciicircum{}(0.5--0.6/CF$_2$)?? ???????? ????????? Cw?? ????????? f\_xy??? ?????? ??????
shoot?? ????????. ??????? ??????? ???????? **?????-??????(lipid-facilitated) ?????? ?????**
(`g\_xy$\cdot$C\_root`, `g\_ph$\cdot$C\_leaf`; B-???????, ???/????? ?????? ????? ?????? ?????)????.

\section{Method}
ORYZA2000 biomass(`oryza\_growth`) + ?????? ??????????, ??? ??????????? free-only(monotone
f\_xy, lipid off)?? lipid-loading(K\_PL-gated g\_xy/g\_ph)?? ????? root/straw/grain BAF???
Yamazaki ???????? ?????????(`validation/longchain\_mechanism.py`). ?????(nC$\geq$10)
straw+grain?? log10 RMSE??? ?? ??????? ??????????, ????????? Cw=C/B?? ???????? B\_root?? ?????
?????????. ?????????? f\_xy=1 ?????? ?????(`validation/refit\_oryza.py`)??? LC1?? ????? ???????
?????. ????? ??????? ??????(Yamazaki)?? ????? in-sample ??????????? ??????????.

Planned approaches:
\begin{itemize}
  \item free-only vs lipid on ORYZA2000 biomass vs Yamazaki
  \item Cw=C/B collapse diagnosis
  \item refit f\_xy=1 ceiling check
\end{itemize}

\section{Hypotheses and findings}

\subsection{Free-anion loading structurally under-predicts long-chain (C10-C12) shoot accumulation; even at the f\_xy=1 ceiling the loaded flux is throttled by the Cw=C/B collapse, so straw/grain cannot reach the observed values.}
\begin{itemize}
  \item Hypothesis id: \texttt{hyp-lc-freethrottle} (confirmatory)
  \item Decision rule (qualitative): free-anion xylem/phloem loading structurally under-predicts long-chain (C10-C12) shoot accumulation -- even at the f\_xy=1 ceiling the flux is throttled by Cw=C/B collapse => support; long chains are reachable by free loading => refute
  \item \textbf{Status: supported} --- confidence strong (graded)
  \item Basis: qualitative judge: Decisive: free-only long-chain straw+grain RMSE 2.026 (\textasciitilde{}100x under), and the ORYZA re-fit hits f\_xy=1/L\_Ph=1 ceilings yet PFDoDA straw stays 14.6 vs 49.8 -- the loaded flux scales with Cw=C/B, which collapses as B grows, so the free-anion shoot cannot reach the long chains. Cause confirmed. Literature corroborates (VERIFIED at source, evi-lc-litread): long-chain PFAS strongly adsorb in roots/soil while short chains are mobile to leaves/grains (newcontam2025); MW is the top predictor of translocation (ML2024).
  \item \textit{Evidence validity:} referent=empirical; data\_source(s)=measured
\end{itemize}

\subsection{A B-independent lipid-facilitated bound-loading term (g\_xy*C\_root, g\_ph*C\_leaf) closes most of the long-chain shoot gap that free loading cannot.}
\begin{itemize}
  \item Hypothesis id: \texttt{hyp-lc-lipidfix} (confirmatory)
  \item Decision rule (qualitative): a B-independent lipid-facilitated bound-loading term (g\_xy*C, g\_ph*C) closes most of the long-chain shoot gap that free loading cannot => support; no improvement => refute
  \item \textbf{Status: supported} --- confidence strong (graded)
  \item Basis: qualitative judge: Decisive: the B-independent lipid term cuts the long-chain straw+grain RMSE 2.026 -> 0.428 (\textasciitilde{}100x -> \textasciitilde{}2.7x) and the whole series 1.035 -> 0.386, recovering PFDA/PFUnDA/PFOS shoot the free model starved. The mechanism is the right direction. SCOPE: in-sample (K\_PL-gated fit), and PFDoDA shoot is still \textasciitilde{}3-4x under -- a residual floor remains. Literature corroborates the MECHANISM, VERIFIED at source (evi-lc-litread): Chen2025 (ES\&T 2025,59,82-91, 10.1021/acs.est.4c06734) -- membrane-water K\_MW rises +0.36/CF2 MONOTONE C4-C16 while HSA (protein) affinity PEAKS at C6-C10, so the lipid (membrane) pool, not protein, carries the longest chains; binding shifts to the membrane with conc.
  \item \textit{Evidence validity:} referent=empirical; data\_source(s)=measured, none
\end{itemize}

\subsection{The lipid mechanism reproduces the long-chain shoot without degrading the root or the short/mid-chain fits (a clean single-mechanism fix).}
\begin{itemize}
  \item Hypothesis id: \texttt{hyp-lc-nocost} (confirmatory)
  \item Decision rule (qualitative): the lipid mechanism reproduces the long-chain shoot WITHOUT degrading the root or the short/mid-chain fits (a clean single-mechanism fix) => support; it trades off root / leaves a residual => refute
  \item \textbf{Status: refuted} --- confidence strong (graded)
  \item Basis: qualitative judge: Decisive: the single-pool lipid term reproduces shoot at the COST of root -- PFUnDA root 20.6->3.9, PFDoDA root 159->4.4 (now under vs obs 69/19.5) -- and PFDoDA shoot is still \textasciitilde{}3-4x under. Not a clean cost-free fix; it needs a 2-pool (free + lipid-bound) split so the bound pool feeds the shoot without draining the root, plus a residual long-chain mechanism.
  \item \textit{Evidence validity:} referent=empirical; data\_source(s)=measured
\end{itemize}

\subsection{A 2-pool root (mobile water+protein pool feeding the xylem + a slow-exchanging lipid/cell-wall bound store holding the measured root burden) reproduces the long-chain shoot while preserving the high measured root, closing the LC3 single-pool tradeoff.}
\begin{itemize}
  \item Hypothesis id: \texttt{hyp-lc-twopool} (confirmatory)
  \item Decision rule (qualitative): a 2-pool root (mobile water+protein pool feeding the xylem + a slow-exchanging lipid/cell-wall bound store holding the measured root burden) reproduces the long-chain shoot WHILE keeping the measured root high, closing the LC3 single-pool tradeoff => support; it cannot match root and shoot simultaneously => refute
  \item \textbf{Status: contested} --- confidence strong (graded)
  \item Basis: qualitative judge: PARTIAL (contested): the 2-pool split RESOLVES the LC3 root tradeoff for mid-long chains -- PFDA(C10) matches root AND shoot simultaneously (3.5/4.2, 5.0/3.5, 4.1/3.4) and PFUnDA(C11) root within \textasciitilde{}2x -- which the single pool could not. But it FAILS for PFDoDA(C12): the mobile pool starves (rm=0.02), so the bound root store (1.2 vs 69) and the shoot both collapse. The PFDoDA residual is an UPTAKE (jR) mass-balance limit, not internal distribution -> a partial fix; PFDoDA needs a different long-chain uptake/irreversible-sorption mechanism.
  \item \textit{Evidence validity:} referent=empirical; data\_source(s)=measured
\end{itemize}

\section{Evidence}
\begin{itemize}
  \item \texttt{evi-lc-2pool-mid} (experiment\_run): point=0.315, finding=validation/twopool\_longchain.py (ORYZA2000): the 2-pool root (mobile + slow lipid/cell-wall bound store; lipid loads fro
  \item \texttt{evi-lc-2pool-pfdoda} (experiment\_run): finding=validation/twopool\_longchain.py: PFDoDA(C12) FAILS -- root collapses to 1.2 vs 69.3 (mobile pool rm=0.02 starves -> boun
  \item \texttt{evi-lc-cost} (experiment\_run): finding=validation/longchain\_mechanism.py: the single-pool lipid term TRADES OFF root for long chains -- PFUnDA root 20.6->3.9 (
  \item \texttt{evi-lc-free} (experiment\_run): point=2.026, finding=validation/longchain\_mechanism.py (ORYZA2000 biomass): free-only (monotone f\_xy) long-chain (nC>=10) straw+grain log10 R
  \item \texttt{evi-lc-lipid} (experiment\_run): point=0.428, finding=validation/longchain\_mechanism.py: the B-independent lipid term (g\_xy*C, g\_ph*C; K\_PL-gated) cuts long-chain straw+grain
  \item \texttt{evi-lc-litread} (literature): finding=5 of 7 corroborating papers obtained out-of-band (paperforge found NO OA PDF for any; 5 supplied manually) and READ to v
  \item \texttt{evi-lit-20260617-153519-b80e9bb0} (literature): finding=\{"acquired": [], "failed": [\{"doi": "10.1021/acs.est.4c06734", "error": "no OA PDF"\}, \{"doi": "10.1021/acs.est.5c11716",
  \item \texttt{evi-pw-decision-20260617-153519-606b27b8} (prior\_work\_decision): finding=searched: DOIs=['10.1021/acs.est.4c06734', '10.1021/acs.est.5c11716', '10.1021/acs.est.7b06128', '10.1021/acsestengg.4c0
\end{itemize}

\section{Pending agent judgments}
These hypotheses use a non-numeric rule and await an in-session agent verdict (no autonomous LLM call):
\begin{itemize}
  \item \texttt{hyp-lc-freethrottle} (qualitative): free-anion xylem/phloem loading structurally under-predicts long-chain (C10-C12) shoot accumulation -- even at the f\_xy=1 ceiling the flux is throttled by Cw=C/B collapse => support; long chains are reachable by free loading => refute
  \item finding: validation/longchain\_mechanism.py (ORYZA2000 biomass): free-only (monotone f\_xy) long-chain (nC>=10) straw+grain log10 RMSE 2.026 (\textasciitilde{}100x); PFDA straw 0.08 vs 3.46, PFDoDA straw 0.33 vs 49.8 / grain 0.10 vs 45.5. refit\_oryza.py: even f\_xy=1/L\_Ph=1 ceilings leave PFDoDA straw 14.6 vs 49.8 -- the Cw=C/B collapse throttles free loading regardless of f\_xy.
  \item \texttt{hyp-lc-lipidfix} (qualitative): a B-independent lipid-facilitated bound-loading term (g\_xy*C, g\_ph*C) closes most of the long-chain shoot gap that free loading cannot => support; no improvement => refute
  \item finding: validation/longchain\_mechanism.py: the B-independent lipid term (g\_xy*C, g\_ph*C; K\_PL-gated) cuts long-chain straw+grain log10 RMSE 2.026 -> 0.428 (\textasciitilde{}100x -> \textasciitilde{}2.7x; whole series 1.035 -> 0.386). PFDA straw 0.08->5.95 (obs 3.46), PFUnDA straw 0.16->11.15 (obs 8.16), PFOS straw 0.17->5.17 (obs 4.35).
\{"acquired": [], "failed": [\{"doi": "10.1021/acs.est.4c06734", "error": "no OA PDF"\}, \{"doi": "10.1021/acs.est.5c11716", "error": "no OA PDF"\}, \{"doi": "10.1021/acs.est.7b06128", "error": "no OA PDF"\}, \{"doi": "10.1021/acsestengg.4c00107", "error": "no OA PDF"\}, \{"doi": "10.48130/newcontam-0025-0007", "error": "no OA PDF"\}, \{"doi": "10.1007/s40726-020-00168-y", "error": "no OA PDF"\}, \{"doi": "10.1139/er-2025-0116", "error": "no OA PDF"\}], "counts": \{"succeeded": 0, "failed": 7\}, "normalization": \{"pdfs": [], "counts": \{"normalized": 0, "already\_extractable": 0, "locked": 0, "error": 0\}\}, "citation\_keys": \{\}, "citation\_key\_collisions": []\}
  \item \texttt{hyp-lc-nocost} (qualitative): the lipid mechanism reproduces the long-chain shoot WITHOUT degrading the root or the short/mid-chain fits (a clean single-mechanism fix) => support; it trades off root / leaves a residual => refute
  \item finding: validation/longchain\_mechanism.py: the single-pool lipid term TRADES OFF root for long chains -- PFUnDA root 20.6->3.9 (obs 19.5), PFDoDA root 159->4.4 (obs 69); and PFDoDA shoot is STILL \textasciitilde{}3-4x under (straw 14.7 vs 49.8). A clean cost-free fix is refuted: it needs a 2-pool (free + bound) split and a residual long-chain mechanism.
  \item \texttt{hyp-lc-twopool} (qualitative): a 2-pool root (mobile water+protein pool feeding the xylem + a slow-exchanging lipid/cell-wall bound store holding the measured root burden) reproduces the long-chain shoot WHILE keeping the measured root high, closing the LC3 single-pool tradeoff => support; it cannot match root and shoot simultaneously => refute
  \item finding: validation/twopool\_longchain.py (ORYZA2000): the 2-pool root (mobile + slow lipid/cell-wall bound store; lipid loads from the mobile pool) matches MID-LONG chains root AND shoot SIMULTANEOUSLY -- PFDA(C10) root 3.5 vs 4.2, straw 5.0 vs 3.5, grain 4.1 vs 3.4; PFUnDA(C11) root 9.8 vs 19.5 (\textasciitilde{}2x), straw 13.4 vs 8.2, grain 3.6 vs 3.1. The single pool could not (LC3 root drain). Shoot RMSE 0.32.
validation/twopool\_longchain.py: PFDoDA(C12) FAILS -- root collapses to 1.2 vs 69.3 (mobile pool rm=0.02 starves -> bound store rb only 1.2; straw 8.6 vs 49.7, grain 26 vs 45.5). The mobile-pool soil uptake (jR) is too small to sustain a high bound root AND feed the shoot: a mass-balance/UPTAKE limit, not internal distribution. The 2-pool does not close the longest chain.
\end{itemize}

\section{References}
\begin{itemize}
  \item \url{https://doi.org/10.1021/acs.est.4c06734}
  \item \url{https://doi.org/10.1021/acs.est.5c11716}
  \item \url{https://doi.org/10.1021/acs.est.7b06128}
  \item \url{https://doi.org/10.1021/acsestengg.4c00107}
  \item \url{https://doi.org/10.48130/newcontam-0025-0007}
  \item \url{https://doi.org/10.1007/s40726-020-00168-y}
  \item \url{https://doi.org/10.1139/er-2025-0116}
\end{itemize}

\end{document}