\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{url}
\title{Yamazaki????? ????????? ?????-?????? ????? ????????????, **???????? ?????? ?????????**(Tang 2026 ???????? TF,}
\author{sci-adk (deterministic render)}
\date{\today}

\begin{document}
\maketitle

\noindent\textit{Draft compiled by sci-adk from Spec \texttt{pfas-rice-oos-multidataset} (v1). Belief state is revisable as Evidence accrues.}

\section{Goal}
Yamazaki????? ????????? ?????-?????? ????? ????????????, **???????? ?????? ?????????**(Tang 2026 ???????? TF,
Kim 2019 ?????? BAF)??? ????????????(monotone f\_xy)$\cdot$???? W2 ???????????? ????????? ?? ??? ?????????,
?? ??????????? ???? ??????????? ????? ?????????(?????? ????? ????????? Li 2025??? ?????????????? group-water$\cdot$
??????????? ????????? inconclusive?? ?????).

\section{Background}
`runs/pfas-rice-oos-lipid`??? ?????-?????? ????? ?????????(`LIPID\_LOADING` ??????? Yamazaki????? ??????,
????? ????????)?? ?????? Tang 2026 ???????? TF?? ??????? ???????? ???????????? ???????(???????????? 1.232 $\rightarrow$
????? 0.516). ?????? ????? **3 ????????(9 ??????)???**????. ?? ?????????? **???? ?????? ??????????? ?????
?????**?????, ??????? Tang ?????? ???????????? ??????????. ???????????? ?? ????? ?????? ??????? ?????????
(Kim 2019 ????? ?????? BAF, `validation/oos\_crossdataset.py`)??, ??????$\cdot$????? ?????????(Li 2025)??
???? ?????.

\section{Method}
?? ????? ??????(monotone `f\_xy\_recommended`=????????????, ???? per-congener `W2fit`, K\_PL-gated
`lipid\_loading=True`)??? **???????? ?????** ?? ?????? ??????????? ??????? ??????? log10 RMSE??? ?????????
(`validation/oos\_multidataset.py` = `oos\_tang\_lipid.py`(Tang) + `oos\_crossdataset.py`(Kim$\cdot$Li)).
1?? ????????? **???????? ?????????(Tang, Kim)**?? ?????; Li??? ??????/group-water/??????????? ?????????
?????????????? ???????? ??????(1?? ????? ?????)????. ??????(Tang 2026$\cdot$Kim 2019$\cdot$Li 2025)?? ????? ???????.

Planned approaches:
\begin{itemize}
  \item ?? ????? ??????(monotone `f\_xy\_recommended`=????????????, ???? per-congener `W2fit`, K\_PL-gated
`lipid\_loading=True`)??? **???????? ?????** ?? ?????? ??????????? ??????? ??????? log10 RMSE??? ?????????
(`validation/oos\_multidataset.py` = `oos\_tang\_lipid.py`(Tang) + `oos\_crossdataset.py`(Kim$\cdot$Li))
  \item 1?? ????????? **???????? ?????????(Tang, Kim)**?? ?????; Li??? ??????/group-water/??????????? ?????????
?????????????? ???????? ??????(1?? ????? ?????)????
  \item ??????(Tang 2026$\cdot$Kim 2019$\cdot$Li 2025)?? ????? ???????.
\end{itemize}

\section{Hypotheses and findings}

\subsection{Yamazaki????? ????????? ?????-?????? ????? ????????????, **???????? ?????? ?????????**(Tang 2026 ???????? TF,
Kim 2019 ?????? BAF)??? ????????????(monotone f\_xy)$\cdot$???? W2 ???????????? ????????? ?? ??? ?????????,
?? ??????????? ???? ??????????? ????? ?????????(?????? ????? ????????? Li 2025??? ?????????????? group-water$\cdot$
??????????? ????????? inconclusive?? ?????).}
\begin{itemize}
  \item Hypothesis id: \texttt{hyp-001} (exploratory)
  \item Decision rule (qualitative): Expert judgment based on evidence
  \item \textbf{Status: supported} --- confidence strong (graded)
  \item Basis: qualitative judge: The lipid-facilitated loading mechanism's out-of-sample generalization is ROBUST across multiple independent datasets, not a Tang artifact: with all params fit on Yamazaki and transferred without refit, lipid wins decisively on BOTH clean datasets (Tang per-organ TF 0.52 vs 1.23; Kim grain 0.48 vs 2.05, reliable 0.20 vs 1.92) and uniquely captures the Kim grain long-chain rise the baselines miss. The only counter-signal is on the pre-registered-as-confounded Li 2025 field data (W2 wins straw/root TF), which does not bear on the clean-dataset claim and where lipid still wins grain/root. Hypothesis supported, with the honest caveats recorded (Li confound; Kim long-chain low-DF). This strengthens hyp-oos-lipid from n=3 (Tang only) to a robust multi-dataset cross-validation.
  \item \textit{Evidence validity:} referent=empirical; data\_source(s)=measured
\end{itemize}

\section{Evidence}
\begin{itemize}
  \item \texttt{evi-oos-multidataset} (experiment\_run): point=0.48, finding=Multi-dataset OUT-OF-SAMPLE robustness of the lipid-facilitated loading mechanism (all transport params fit on Yamazaki,
\end{itemize}

\section{Pending agent judgments}
These hypotheses use a non-numeric rule and await an in-session agent verdict (no autonomous LLM call):
\begin{itemize}
  \item \texttt{hyp-001} (qualitative): Expert judgment based on evidence
  \item finding: Multi-dataset OUT-OF-SAMPLE robustness of the lipid-facilitated loading mechanism (all transport params fit on Yamazaki, transferred WITHOUT refit). On the two CLEAN independent datasets the lipid mechanism wins decisively over both the free-anion/monotone f\_xy and the saturated W2 fit: (1) Tang 2026 per-organ TF (dw) log10 RMSE lipid 0.52 vs free-anion/monotone 1.23; (2) Kim 2019 brown-rice grain BAF (porewater, excl PFOA used in the L\_Ph fit) lipid 0.48 vs monotone 2.05 vs W2 1.07, and on the reliable (DF>=15\%) congeners lipid 0.20 vs monotone 1.92 vs W2 1.44 -- lipid is the only variant that captures the grain chain-length pattern incl. the long-chain RISE that monotone/W2 structurally miss (mono gives O(0.05) for PFUnDA/PFDoDA obs \textasciitilde{}33-35). The generalization is therefore NOT a Tang artifact -- it holds across two independent datasets (Korean field grain + Chinese pot per-organ). Honest limits (pre-registered): (a) Li 2025 is field/group-water/surface-confounded and inconclusive as expected -- lipid wins grain/root TF (0.72 vs mono 1.15, W2 1.47) but W2 wins straw/root TF (0.33 vs lipid 0.57, mono 1.03); (b) Kim long chains are low detection-frequency (DF 3-13\%) so the headline excl-PFOA score leans on them. Primary verdict rests on the clean datasets (Tang, Kim).
\end{itemize}

\section{References}
No literature cited.

\end{document}